### 2022-02-01 Manuscript/Jacobsville_manuscript.tex

{\Large \textbf{Laurentia's continued drifting into the Neoproterozoic: new paleomagnetic pole from the upper Jacobsville sandstone}}
\\
\singlespacing
 Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}\\
\bigskip
\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA\\

@@

\section*{ABSTRACT}
New paleomagnetic pole developed from red beds in the upper section of the ca. 990 Ma Jacobsville Sandstone positions Laurentia in the southern hemisphere. The new data suggest Laurentia's continued southerly motion during post-Midcontinent Rift magmatism. This paleomagnetic pole from the intracontinental sedimentary package is inconsistent with Laurentia to have experienced 4000 km of shortening associated with the Grenville Orogeny.

@@

>>>>

{\Large \textbf{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic pole from the Jacobsville sandstone}}
\\
\singlespacing
 Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}\\
\bigskip
\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA\\

@@

\section*{ABSTRACT}

The ca. 1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments in the interior of Laurentia while the continent traveled rapidly from high to low latitudes. Coeval with the cessation of rifting and decrease in the speed of plate motion is the ca. 1090-980 Ma Grenvillian Orogeny whose far-field stress is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation. A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in a gap of record for the position and orientation of Laurentia for nearly 100 Myr. We develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US. High-resolution thermal demagnetization experiments together with successful fold tests resolve primary magnetic remanence which show shallow inclinations despite being  flattened. Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south to the Keweenawan Track ca. 990 Ma, confirming continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic. The new pole from the interior of Laurentia indicate the Grenville Loop poles are younger than previously thought.

@@

Paleomagnetic data for the Jacobsville have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig. 7; Table 4). As a result, its age has been interpolated to be latest Mesoproterozoic (e.g., 1020 Ma in Li et al., 2008) such that it was deposited prior to the ca. 1000 Ma Grenville loop poles. However, this age assignment has recently been questioned on the basis of laser-ablation–inductively cou- pled plasma–mass spectrometry (LA-ICP-MS) U-Pb detrital zircon dates from Jacobsville Sandstone samples. Malone et al. (2016) interpreted a maximum depositional age of 959 ± 19 Ma based on the weighted mean of the four youngest grains from 2050 LA-ICP-MS zircon dates. A similar age was proposed on the basis of the youngest LA-ICP-MS U-Pb dates in the study of Craddock et al. (2013). This young maximum interpreted age is intriguing because it suggests the unconformity above the Freda Sandstone spans at least 100 m.y. and was followed by continued sedi- mentation in the region long after rift activity. Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that is consistent with deposition of the Jacobsville Sand- stone in the late Mesoproterozoic. Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the pos- sibility of outlier dates in such a large data set, it would be beneficial for the youngest Jacobsville zircons to be analyzed using CA-ID-TIMS to obtain more precise constraints.
